
\newpage

# Systematic sampling: 3 x 3 km grid {#sys3}

`r ifelse(book_format == "html" & with_watermark, "<div class='watermark'>DRAFT</div>", "")`

This section is optional. If you wish to practice more grid creation commands, or if you are curious to know how many forest plots we would have with a 3 x 3 km grid, complete the exercises below, otherwise you can skip directly to the next session.

**In the following console boxes, create `sf_point3`, `sf_plot3`, `nplot3` and `nplot3_total` for creating a 3 x 3 km grid (we skip the grid itself to start directly with the points):** 

- Points with the option `what = "center"`.

```{r points3-cons, eval=F, class.source="Rchunk-cons"}

## Console
##

```

```{r points3-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

## Solution
sf_points3 <- st_make_grid(sf_lc, cellsize = c(3000, 3000), what = "centers", offset = offset) %>%
  st_intersection(sf_admin) %>%
  st_as_sf()

```


- Plots by joining the land cover shapefile `sf_lc`.

```{r plot3-cons, eval=F, class.source="Rchunk-cons"}

## Console
##

```

```{r plot3-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

## Solution
sf_plot3 <- sf_points3 %>%
  st_join(sf_lc) %>%
  mutate(lc = fct_reorder(lc, lc_id)) %>%
  filter(!is.na(lc))

```

\  

- Make a map with the plots over the land cover and use our color palette `pal` for ```r newland_name``` land covers.

```{r ggplot3-cons, eval=F, class.source="Rchunk-cons"}

## Console
##

```

```{r ggplot3-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

## Solution
ggplot() +
  geom_sf(data = sf_lc, aes(fill = lc), color = NA) +
  geom_sf(data = sf_plot3, aes(fill = lc), shape = 21) +
  geom_sf(data = sf_admin, fill = NA) +
  scale_fill_manual(values = pal) +
  labs(fill = "", color = "") +
  theme_void()

```

\  

- Calculate the number of plots per land cover class using the `group_by()` and `summarise()` functions.

```{r nplot3-cons, eval=F, class.source="Rchunk-cons"}

## Console
##

```

```{r nplot3-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

## Solution
nplot3 <- sf_plot3 %>%
  as_tibble() %>%
  group_by(lc) %>%
  summarise(n = n())
nplot3

```

\  

- Finally calculate the total number of plots in forest land.

```{r nplot3tot-cons, eval=F, class.source="Rchunk-cons"}

## Console

```

```{r nplot3tot-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

## Solution
nplot3_total <- nplot3 %>%
  filter(!(lc %in% c("WA", "NF"))) %>%
  summarise(n = sum(n))
nplot3_total

```


