


### The 3 km grid {-}

**You turn! Prepare `sf_point3`, `sf_plot3`, `nplot3` and `nplot3_total` for creating a 3 km grid.** 

Create directly:

- Points with the option `what = "center"`, this time the `cellsize` argument will be 3000 m.

```{r points3-cons, eval=F, class.source="Rchunk-cons"}

## Console
sf_points3 <- st_make_grid(sf_lc, cellsize = c(___, ___), what = "centers", offset = offset) %>%
  st_intersection(sf_admin) %>%
  st_as_sf()

```

```{r points3-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

sf_points3 <- st_make_grid(sf_lc, cellsize = c(3000, 3000), what = "centers", offset = offset) %>%
  st_intersection(sf_admin) %>%
  st_as_sf()

```


- Plots by joining the land cover shapefile `sf_lc`.

```{r plot3-cons, eval=F, class.source="Rchunk-cons"}

## Console
sf_plot3 <- sf_points3 %>%
  st_join(___) %>%
  mutate(lc = fct_reorder(lc, lc_id)) %>%
  filter(!is.na(lc))

```

```{r plot3-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

sf_plot3 <- sf_points3 %>%
  st_join(sf_lc) %>%
  mutate(lc = fct_reorder(lc, lc_id)) %>%
  filter(!is.na(lc))

```

\  

- A map with the plots over the land cover (remember we have a custom color palette `pal` for ```r newland_name``` land covers).

```{r ggplot3-cons, eval=F, class.source="Rchunk-cons"}

## Console
ggplot() +
  geom_sf(data = ___, aes(fill = lc), color = NA) +
  geom_sf(data = ___, aes(fill = lc), shape = 21) +
  scale_fill_manual(values = pal) +
  labs(fill = "", color = "") +
  theme_void()

```

```{r ggplot3-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

ggplot() +
  geom_sf(data = sf_lc, aes(fill = lc), color = NA) +
  geom_sf(data = sf_plot3, aes(fill = lc), shape = 21) +
  scale_fill_manual(values = pal) +
  labs(fill = "", color = "") +
  theme_void()

```

\  

- Calculations of the number of plots per land cover class.

```{r nplot3-cons, eval=F, class.source="Rchunk-cons"}

## Console
nplot3 <- ___ %>%
  as_tibble() %>%
  group_by(lc) %>%
  summarise(n = ___)
nplot3 

```

```{r nplot3-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

nplot3 <- sf_plot3 %>%
  as_tibble() %>%
  group_by(lc) %>%
  summarise(n = n())
nplot3 

```


- And the total number of plot in forests.

```{r nplot3tot-cons, eval=F, class.source="Rchunk-cons"}

## Console
nplot3_total <- ___ %>%
  filter(!(lc %in% c("WA", "NF"))) %>%
  summarise(___)
nplot3_total

```

```{r nplot3tot-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

nplot3_total <- nplot3 %>%
  filter(!(lc %in% c("WA", "NF"))) %>%
  summarise(n = sum(n))
nplot3_total

```

\  

Good job! The 3 km grid looks like exactly what we need. Let's still design a 2 km grid to see if we could afford it.

\  
