
\newpage

# Random Sampling {ran}

`r ifelse(book_format == "html" & with_watermark, "<div class='watermark'>DRAFT</div>", "")`

Complete random sampling designs have a number of disadvantages and are generally not recommended at a national scale. We will still relocate our ```r nplot3_total$n``` plots randomly to compare different sampling designs. 


## Demo {-}

The equivalent of `st_make_grid()` for random sampling is `st_sample()`. Every time `st_sample()` is used a new set of random locations are created. To get the same random allocation every time we need to set the random selection with a seed and the function `set.seed()`. In the example we will use `set.seed(10)` but feel free to change or remove at will the chose seed number in the practice below.

\  

::::::{.infobox data-latex=""}

As the plot locations are completely random, we have no control over how many plots are going to land on forest versus non-forest land. We can apply two possible strategies to reach our target number of plots:

1. We filter out non-forest polygons and make the sample size equal to the total number of plots in forest lands from our 4 x 4 km sampling grid `nplot4_total$n`.
1. We make the sampling size equal to the original number of points in a 4 x 4 km grid `nrow(sf_points4)`

We can anticipate that the second solution is better if you are also interested in non-forest land, although this is no guarantee you will have enough plots in each land cover. We will favor option 1 as our primary target is forest carbon stock.

As `sf` objects behave similarly to tibbles, the tidyverse function `filter()` will allow us to prepare a forest only shapefile.

```{r forest}

sf_forest <- sf_lc %>%
  filter(lc %in% c("EV", "MD", "DD", "WL", "MG"))

```

::::::

\  

We can now create the random points based on the `sf_forest` canvas.

```{r simp-sample}

set.seed(10)
sf_points_rd <- st_sample(x = sf_forest, size = nplot4_total$n) %>%
  st_as_sf()

```

\  

We can then extract the land cover at the plots' location...

```{r simp-extract}

sf_plot_rd <- sf_points_rd %>%
  st_join(sf_lc) %>%
  mutate(lc = fct_reorder(lc, lc_id)) %>%
  filter(!is.na(lc))

```

... And map them.

```{r simp-map}

ggplot() +
  geom_sf(data = sf_lc, aes(fill = lc), color = NA) +
  geom_sf(data = sf_plot_rd, aes(fill = lc), shape = 21) +
  geom_sf(data = sf_admin, fill = NA) +
  scale_fill_manual(values = pal) +
  labs(fill = "", color = "") +
  theme_void()

```

\  

If at a first glance the plot distribution shows a decent coverage of the new land, it is interesting to calculate the number of plots per land cover category.

```{r simp_nplot}

nplot_rd <- sf_plot_rd %>%
  as_tibble() %>%
  group_by(lc) %>%
  summarise(n = n())
nplot_rd

```

\  

## Practice {-}

**Your turn! In the console below, test different seed numbers and visualize the results.**

- Create the points, choose whatever seed number you want:

```{r ranpoints-cons, eval=F, class.source="Rchunk-cons"}

## Console
set.seed(___)
sf_points_rd2 <- st_sample(x = ___, size = ___) %>%
  st_as_sf()

```

```{r ranpoints-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

set.seed(100)
sf_points_rd2 <- st_sample(x = sf_lc, size = nplot3_total$n) %>%
  st_as_sf()

```

- Add the land cover information:

```{r ranplot-cons, eval=F, class.source="Rchunk-cons"}

## Console
sf_plot_rd2 <- sf_points_rd2 %>%
  st_join(___) %>%
  mutate(lc = fct_reorder(lc, ___)) %>%
  filter(___)

```

```{r ranplot-sol, class.source="Rchunk-sol", Rout="Rout-sol"}

sf_plot_rd2 <- sf_points_rd2 %>%
  st_join(sf_lc) %>%
  mutate(lc = fct_reorder(lc, lc_id)) %>%
  filter(!is.na(lc))

```

\  

- Show the plot locations on a map:

```{r ranmap-cons, eval=F, class.source="Rchunk-cons"}

## Console
ggplot() +
  geom_sf(data = sf_lc, aes(fill = lc), color = NA) +
  geom_sf(data = ___, aes(fill = ___), shape = 21) +
  geom_sf(data = sf_admin, fill = NA) +
  scale_fill_manual(values = pal) +
  labs(___, ___) +
  theme_void()

```

```{r ranmap-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

ggplot() +
  geom_sf(data = sf_lc, aes(fill = lc), color = NA) +
  geom_sf(data = sf_plot_rd2, aes(fill = lc), shape = 21) +
  geom_sf(data = sf_admin, fill = NA) +
  scale_fill_manual(values = pal) +
  labs(fill = "", color = "") +
  theme_void()

```

- Calculate the number of plots per land cover type:

```{r rancount-cons, eval=F, class.source="Rchunk-cons"}

## Console
nplot_rd2 <- ___ %>%
  as_tibble() %>%
  group_by(___) %>%
  summarise(n = n())
nplot_rd2

```

```{r rancount-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

nplot_rd2 <- sf_plot_rd2 %>%
  as_tibble() %>%
  group_by(lc) %>%
  summarise(n = n())
nplot_rd2

```

\  

Well done! We will keep these results for later and move on with Uneven Systematic Sampling.

\  

