

# Uneven Systematic Sampling {uneven}

In the NFI technical module 3 on sampling, we have seen that systematic sampling are popular for their increased precision and low auto-correlation compared to simple random sampling. We have also seen that their main caveat is that a systematic sample is based on only one randomization. One way to avoid this limitation is to create an <txt-green>uneven systematic sampling</txt-green>. 

An uneven systematic sample starts with the same grid as a systematic sample, but instead of placing plots at the center or corner of the grid, we allocate them randomly inside each square of the grid. 

\  

## Demo {-}

We will demonstrate the R commands for the 5 x 5 km grid and let you practice with the 4 x 4 km grid that we have selected for ```r newland_name```'s NFI. For making random points within a grid, we will use the function `st_sample()`, but this time we will apply it to the grid polygons and the size of the sample is 1 but repeated for each points in the grid.

```{r uneven5}

set.seed(10)
sf_points_uneven <- st_sample(x = sf_grid5, size = rep(1, nrow(sf_points5)))

```

Same as before, we can overlay the results with the land cover shapefile to see how many plots would fall on forests.

```{r uneven5-plot}

sf_plot_uneven <- sf_points_uneven %>%
  st_as_sf() %>%
  st_join(sf_lc) %>%
  mutate(lc = fct_reorder(lc, lc_id)) %>%
  filter(!is.na(lc))

```

Then a quick map is always a good idea to see if everything looks at it should. In this case, we should not see more than one pot in each grid square.

```{r uneven5-map}

ggplot() +
  geom_sf(data = sf_lc, aes(fill = lc), color = NA) +
  geom_sf(data = sf_grid5, fill = NA, col = "red") +
  geom_sf(data = sf_plot_uneven, aes(fill = lc), shape = 21) +
  geom_sf(data = sf_admin, fill = NA) +
  scale_fill_manual(values = pal) +
  labs(fill = "", color = "") +
  theme_void()

```


**Your turn! Create an uneven systematic sample based on the 4 x 4 km grid.**

- First you need to create the 4 x 4 km grid. Have a look at the section \@ref{sysdemo} for code block if you don't remember the function inputs, and adpat it to the 4 x 4 km grid.

```{r grid4-cons, eval=F, class.source="Rchunk-cons"}

# Console
sf_grid4 <- st_make_grid(sf_lc, cellsize = c(___, ___), what = ___, offset = offset) %>%
  st_intersection(sf_admin)

```

```{r grid4-sol, class.source="Rchunk-sol", Rout = "Rout-sol"}

sf_grid4 <- st_make_grid(sf_lc, cellsize = c(4000, 4000), what = "polygons", offset = offset) %>%
  st_intersection(sf_admin)

```


- Create the points. We have just created the grids canvas `sf_grid4` and we have already created the grid points `sf_points4` in the section \@ref(sys4). Keep the seed 10 to have comparable results with all the users of this lesson. 

NB: This operation is calculation intensive and may take a few minutes to complete.

```{r unevenpoints-cons, eval=F, class.source="Rchunk-cons"}

## Console
set.seed(10)
sf_points4_uneven <- st_sample(x = ___, size = ___) %>%
  st_as_sf()

```

```{r unevenpoints-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

set.seed(10)
sf_points4_uneven <- st_sample(x = sf_grid4, size = rep(1, nrow(sf_points4))) %>%
  st_as_sf()

```

- Add the land cover information:

```{r unevenplot-cons, eval=F, class.source="Rchunk-cons"}

## Console
sf_plot4_uneven <- ___ %>%
  ___(___) %>%
  ___(lc = fct_reorder(lc, ___)) %>%
  ___(___)

```

```{r unevenplot-sol, class.source="Rchunk-sol", Rout="Rout-sol"}

sf_plot4_uneven <- sf_points4_uneven %>%
  st_join(sf_lc) %>%
  mutate(lc = fct_reorder(lc, lc_id)) %>%
  filter(!is.na(lc))

```

\  

- Show the plot locations on a map with a `ggplot()`:

```{r unevenmap-cons, eval=F, class.source="Rchunk-cons"}

## Console
##

```

```{r unevenmap-cons2, eval=F, class.source="Rchunk-cons"}

## Tip
ggplot() +
  geom_sf(data = sf_lc, aes(fill = lc), color = NA) +
  geom_sf(data = ___, fill = NA, col = "red") +
  geom_sf(data = ___, aes(fill = lc), shape = 21) +
  geom_sf(data = ___, fill = NA) +
  scale_fill_manual(values = ___) +
  ___(fill = "", color = "") +
  theme_void()

```

```{r unevenmap-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

## Solution
ggplot() +
  geom_sf(data = sf_lc, aes(fill = lc), color = NA) +
  geom_sf(data = sf_grid4, fill = NA, col = "red") +
  geom_sf(data = sf_plot4_uneven, aes(fill = lc), shape = 21) +
  geom_sf(data = sf_admin, fill = NA) +
  scale_fill_manual(values = pal) +
  labs(fill = "", color = "") +
  theme_void()

```

\  

- Calculate the number of plots per land cover type:

```{r unevencount-cons, eval=F, class.source="Rchunk-cons"}

## Console
##

```


```{r unevencount-cons2, eval=F, class.source="Rchunk-cons"}

## Tips
nplot4_uneven <- ___ %>%
  as_tibble() %>%
  group_by(___) %>%
  summarise(n = n())
nplot4_uneven

```

```{r unevencount-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

## Solution
nplot4_uneven <- sf_plot4_uneven %>%
  as_tibble() %>%
  group_by(lc) %>%
  summarise(n = n())
nplot4_uneven

```

\  

Well done! We have one of the best possible simple sampling design, combining strengths of both systematic and random sampling.


```{r, echo=F}

simple_res <- area_lc %>%
  left_join(nplot4 %>% rename(n_sys4 = n), by = "lc") %>%
  left_join(nplot_rd %>% rename(n_rd = n), by = "lc") %>%
  left_join(nplot4_uneven %>% rename(n_uneven4 = n), by = "lc") 

simple_tot <- simple_res %>%
  summarise(across(where(is.numeric), ~ sum(., na.rm = TRUE))) %>%
  mutate(lc = "Total") %>% 
  select(lc, everything())
  
simple_perc <- simple_res %>%
  mutate(
    area_p = area_ha / simple_tot$area_ha,
    
  )

  

```
\  

